---
title: "runningSLIDE"
output: rmarkdown::html_vignette
date: "2023-11-01"
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Install SLIDEHelper
SLIDEHelper will install the latest versions of SLIDE and Essential Regression
```{r libinstall, eval = F, tidy=TRUE}
library(devtools)
devtools::install_github(repo = "Hanxi-002/SLIDEHelper", force = F,
                         dependencies = T)
```

Note that this package will be moved to `jishnu-lab/SLIDE` in future versions.
```{r imports, eval = T, message = F}
library(SLIDEHelper)
library(tidyverse)
library(doParallel)
```

### Run Essential Regression
```{r essreg, eval = F}
#!/usr/bin/env Rscript
args = commandArgs(trailingOnly=TRUE)

library(SLIDEHelper)
library(tidyverse)
library(doParallel)

cores <-  as.numeric(Sys.getenv('SLURM_CPUS_PER_TASK', unset=NA))
if(is.na(cores)) cores <- detectCores()
registerDoParallel(cores)
cat('number of cores using', cores, '. . .\n')

EssReg::pipelineER3(yaml_path)
```

You can get more detailed error messages by running `EssReg::pipelineER3(yaml_path)` within the `withCallingHandlers` R function like so:
```{r getER_errors, eval = F}
withCallingHandlers(
  EssReg::pipelineER3(yaml_path), 
  error = function(e) { write.to.log(sys.calls()) })
```

### Run SLIDE

##### Load ER results
After running ER, you can load your data for SLIDE using the yaml file:
```{r findER_Results, eval = F}
yaml_input = yaml::yaml.load_file(yaml_path)

# this looks for the ER results from pipeline3 that begins with name 'final_delta...' 
er_results_path = list.files(er_path, pattern = "final_delta_", full.names = T)
er_results = readRDS(er_results_path)

x_path = yaml_input$x_path
y_path = yaml_input$y_path

x_mat = as.matrix(read.csv(x_path, row.names = 1, check.names = F))
y_mat = as.matrix(read.csv(y_path, row.names = 1, check.names = F))

```

You can view example data (input x and y, z_matrix and SLIDE results) from the SLIDE manuscript:
```{r SLIDEHelper_example_data, eval = T}
data(SSc_data)

# data stored in a list
names(SSc_data)
```

#### Run SLIDE
```{r SLIDEHelper, eval = F}
z_mat = EssReg::predZ(x_mat, er_results)
  colnames(z_mat) = paste0("Z", c(1:ncol(z_mat)))
  rownames(z_mat) = rownames(y_mat)

# save this z matrix with ER results  
write.csv(z_mat, paste0(er_path, "/z_mat.csv"))

SLIDE_res <- runSLIDE(y_path = y_path,
                        z_path = NULL,
                        z_matrix = z_mat,
                        er_path = er_results_path,
                        do_interacts = TRUE,
                        spec = 0.2,
                        niter = 1000)
```

#### Get significant latent factor features
```{r SLIDEfeat, eval = F}
num_top_feats <- 10
condition <- "auc"

SLIDE_res <- SLIDEHelper::GetTopFeatures(x_path, y_path, er_results_path, out_path = er_path, SLIDE_res, num_top_feats = num_top_feats, condition)

saveRDS(SLIDE_res, paste0(er_path, "/SLIDE_res.RDS"))
```

#### Plot significant marginal and interaction latent factors
```{r plotSigs, eval = F}
# plotSigGenes(SLIDE_res, out_path = er_path, plot_interaction = TRUE)
plotSigGenes(SSC_data$SLIDE_res, out_path = ".", plot_interaction = TRUE)

```

#### Plot SLIDE performance
```{r calcperf, eval = F}
#the SLIDE_res has to be the output from GetTopFeatures
# CalcControlPerformance(z_matrix = z_mat, y_path, SLIDE_res, niter = 1000, condition, out_path = er_path)
CalcControlPerformance(z_matrix = SSc_data$z_mat, y_path, SLIDE_res, niter = 1000, condition, out_path = er_path)

```


